---
title: "Covid91 vaccine study Final2023"
author: "Lisa Davis"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalData  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$: The covid-19 vaccine (drug) is not effective among the subgroup users.
$H_A$: The covid-19 vaccine (drug) is effective among non subgroup users.

## Males

### Introduction

$H_0$: The vaccine is not effective in treating covid-19 in males.

$H_A$: The vaccine is effective in treating covid-19 in males.

The data used for this clinical test, came from a final data file provided by professor Brownlow. 


### Methods

The response method used is Cat~Cat. Data used was based off data from the final data provided, we also subset the data to include males. We will be using bar chart, numerical results, and inferential results for the graphical display.

### Results

```{r}
maleData = filter (FinalData, sex == "M")
```

#### Descriptive results

##### Graphical results

```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")
```

The graph shows, among the males who got covid-19, about 30% of the vaccinated group had symptoms, and about 70% of the  placebo group had symptoms. The graph also shows that there was a slight difference in the all male group whether vaccinated or not vaccinated there was a slight to almost no difference between the two

##### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

The data shows that there was a 2 times difference for the males who took the vaccine and got covid-19. The males who took the placebo had 70% chance of not being infected, and 30% less likely, for the male who took the vaccine. Among the males who were ok and took the vaccine, there was no difference.

##### Inferencial results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Based on the chisq-test, the p value is 0.0000000000001098 is so small, so we rejected the null hypothesis.


### Discussion and conclusion

The clinical test concludes that the vaccine is not effective in treating covid-19 among males.
The p value is 0.0000000000001098 is so small, so we rejected the null hypothesis.
The Fisher Exact test p value is 0.00000000000004613, which means there is a difference in getting covid-19 being a male.



## Females

### Introduction

$H_0$: The vaccine is not effective in treating covid-19 in females.

$H_A$: The vaccine is effective in treating covid-19 in females.

The data used for this clinical test, came from a final data file provided by professor Brownlow. 


### Methods

The response method used is Cat~Cat. Data used was based off data from the final data provided, we also subset the data to include females. We will be using bar chart, numerical results, and inferential results for the graphical display.


### Results

```{r}
femaleData = filter(FinalData, sex == "F")
```

#### Descriptive results

##### Graphical results

```{r}
barchartGC(~infected+treatment,data=femaleData)
barchartGC(~infected+treatment,data=femaleData, type="percent")
```

The graph shows, among the females who got covid-19, about 40% of the vaccinated group had symptoms, and about 60% of the  placebo group had symptoms. The graph also shows that there was a slight difference in the all females group whether vaccinated or not vaccinated there was a slight to almost no difference between the two.

##### Numerical results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```
The data shows among the females group that were infected by covid-19, those were unvaccinated, 60% of females experienced symptoms, while only 40% of female group who were vaccinated experienced symptoms.The data also shows that within the female placebo group, 2.04% were infected with covid-19, while only 1.37% of the female vaccinated group contracted covid-19.

##### Inferencial results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Based on this chisq test, the p-value is 0.0002732,is so small, we rejected the null hypothesis.


### Discussion and conclusion



## LGBTQ

### Introduction

$H_0$: The vaccine is not effective in preventing covid-19 in LGBTQ.

$H_A$: The vaccine is effective in preventing covid-19 in LGBTQ.

The data used for this clinical test, came from a final data file provided by professor Brownlow. 

### Methods

The response method used is Cat~Cat. Data used was based off data from the final data provided, we also subset the data to include LGBTQ. We will be using bar chart, numerical results, and inferential results for the graphical display.


### Results

```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```


#### Descriptive results


##### Graphical results

```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```

The graph shows, among the LGBTQ who got covid-19, about 80% of the vaccinated group had symptoms, and about 1% of the  placebo group had symptoms. The graph also shows that there was a slight difference in the all LGBTQ group whether vaccinated or not vaccinated there was a slight to almost no difference between the two.

##### Numerical results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData) 
rowPerc(table3)
colPerc(table3)
```

The numerical data shows that within the LGBTQ placebo group, they experienced 15.38% of symptoms, while the vaccinated LGBTQ group experienced 84.62% of syptoms. The data also shows that 0.66% of the LGBTQ group who had the placebo contracted covid-19, while 3.54% of the vaccinated group contracted covid-19.



##### Inferencial results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

Based on this chisq test, the p-value 0.0000006668 is so small, we reject the null hypothesis.


### Discussion and conclusion

The chisq test shows that the p-value for the LGBTQ group is 0.0000006668 showing that there is an effect between the COVID vaccine and the LGBTQ subgroup, concluding that we are rejecting the null hypothesis.  

In the fisher exact test, the percentage odds ratio is about 0.18%, showing that the odds are about 0.18 times lower in one variable group compared to the other, showing a difference between the two groups, which is another example of rejecting the null hypothesis in this group.

## Druggies

### Introduction

$H_0$: The vaccine is not effective in preventing covid-19 in Druggies.

$H_A$: The vaccine is effective in preventing covid-19 in Druggies.

The data used for this clinical test, came from a final data file provided by professor Brownlow. 

### Methods

The response method used is Cat~Cat. Data used was based off data from the final data provided, we also subset the data to include Druggies. We will be using bar chart, numerical results, and inferential results for the graphical display.


### Results

```{r}
DrugUserData = filter(FinalData, DrugUser == "yes")
```


#### Descriptive results


##### Graphical results

```{r}
barchartGC(~infected+treatment,data=DrugUserData)
barchartGC(~infected+treatment,data=DrugUserData, type="percent")
```

The graph shows among the DrugUser who got covid-19, about 80% of the vaccinated group had symptoms, and about 15% of the  placebo group had symptoms. The graph also shows that there was a slight difference in the all DrugUser group whether vaccinated or not vaccinated there was a slight to almost no difference between the two.


##### Numerical results

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUserData) 
rowPerc(table4)
colPerc(table4)
```

The data shows that 11.67% of the placebo druggie group experienced covid-19 symptoms when they were infected by covid-19, while 88.33% of vaccinated DrugUsers experienced symptoms. The data also shows .4% of the druggie placebo group got infected with covid-19, while 3.06% of the vaccinated group got infected with covid-19.

##### Inferencial results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

Based on this chisq test, the p-value 0.00000002842 is so small, we reject the null hypothesis.


### Discussion and conclusion

The p value of .000000002842 indicates a high significant association between the variables which caused us to reject the null hypothesis.

The fisher exact test showed a p-value of 0.00000232, which is showing that there is a  odds ratio of about 0.12%, suggesting that the odds of this occurring are lower in one group compared to the other, the 95% confidence interval supports this.


# Overall Results and Conclusions

In conclusion, the effectiveness of the COVID vaccine varies between subgroups. Starting with the all male group, we are rejecting the null hypothesis.

Based on the bar chart, numerical graph, and the chisq and fisher tests, we can conclude that men who received the COVID vaccine and contracted it ultimately experience less symptoms, than those who received the placebo. The men that did not get vaccinated and did not contract COVID showed little to no difference between each other. In the group that did receive the placebo had an infection rate of 2.55%, and those of them who got the vaccine had an infection rate of 1.15%, showing that the vaccine is effective in men, therefore we are rejecting the null hypothesis for the male group.


The all female group, we are also rejecting the null hypothesis.

Based on all of the tests we have run, we can conclude that women who received the COVID vaccine and contracted it also experienced less symptoms than the women who received the placebo. The female group that did not get vaccinated and did not contract COVID also showed little to no difference compared to each other. Within the group that did receive the placebo, they had an infection rate of 2.04%. The group that received the vaccine had an infection rate of 1.37%, concluding that the vaccine is effective in women, therefore rejecting the null hypothesis for the female subgroup.

Comparing the male and female subgroups, women had a smaller percentage of infection with the COVID vaccine and placebo than then men by 0.22% and 0.51%. It is not a big difference, but so far within the data analysis women experience less symptoms with the COVID vaccine.

The LGBTQ subgroup, we are also rejecting the null hypothesis.

Based on our tests, we can conclude that the COVID vaccine is effective in this subgroup, but negatively. We found out that the LGBTQ people who received the COVID vaccine and contracted it experienced significant symptoms than those who received the placebo. The LGBTQ group that did not get vaccinated and did not contract COVID were almost exactly the same, showing no difference. In the LGBTQ group that did receive the placebo had an infection rate of 0.66% and those who did receive the vaccine had an infection rate of 3.54%, a very significant difference between the two, therefore rejecting the null hypothesis since the vaccine had an effect on the group.

The drug user group, we are also rejecting the null hypothesis.

Finally, for the drug user group, they received the same result as the LGBTQ subgroup. We are rejecting the null hypothesis due to the drug users experiencing significantly more symptoms than the placebo group, showing that the COVID vaccine did have an effect, but not a positive one In the group of drug users that were vaccinated and contracted COVID, they had a 3.06% infection rate. The placebo group had a 0.4% infection rate showing a very big difference between the two and how much symptoms each group got. 

Comparing all of the tests from most positive COVID affects to the more negative COVID affects, the female had the best results with the COVID vaccine, while the LGBTQ subgroup had it the worst, experiencing the most symptoms while having COVID even with the vaccine. Males had almost the same results as the females, just a little less and drug users had fewer symptoms by only 0.48%. I would recommend that only males and females that are not drug users or LGBTQ to get the COVID vaccine since the effect was more positive than it was for the other two subgroups. 

