---
title: "Covid91 vaccine study Final2023"
author: "Lisa Davis"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalData  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$: The covid-19 vaccine (drug) is not effective among the subgroup users.
$H_A$: The covid-19 vaccine (drug) is effective among non subgroup users.

## Males

### Introduction

$H_0$: The vaccine is not effective in preventing covid-19 in males.

$H_A$: The vaccine is effective in preventing covid-19 in males.

The data used for this clinical test, came from a final data file provided by professor Brownlow. 


### Methods

The response method used is Cat~Cat. Data used was based off data from the final data provided, we also subset the data to include males. We will be using bar chart, numerical results, and inferential results for the graphical display.

### Results

```{r}
maleData = filter (FinalData, sex == "M")
```

#### Descriptive results

##### Graphical results

```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")
```

The graph shows, among the males who got covid-19, about 30% got the vaccine, and about 70% got placebo. The graph also shows that there was a slight difference in the males who received the vaccine, and were ok.

##### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

The data shows that there was a 2 times difference for the males who took the vaccine and got covid-19. The males who took the placebo had 70% chance of not being infected, and 30% less likely, for the male who took the vaccine. Among the males who were ok and took the vaccine, there was no difference.

##### Inferencial results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value is 0.0000000000001098 is so small, so we failed to reject the null hypothesis.


### Discussion and conclusion

The clinical test concludes that the vaccine is not effective in preventing covid-19 among males.
The p value is 0.0000000000001098 is so small, so we failed to reject the null hypothesis.
The Fisher Exact test p value is 0.00000000000004613, which means there is a difference in getting covid-19 being a male.



## Females

### Introduction

### Methods

### Results

#### Descriptive results

##### Graphical results

##### Numerical results

##### Inferencial results

### Discussion and conclusion



## LGBTQ

### Introduction

### Methods

### Results

#### Descriptive results

##### Graphical results

##### Numerical results

##### Inferencial results

### Discussion and conclusion



## Druggies

### Introduction

### Methods

### Results

#### Descriptive results

##### Graphical results

##### Numerical results

##### Inferencial results

### Discussion and conclusion



# Overall Results and Conclusions

### Introduction

### Methods

### Results

#### Descriptive results

##### Graphical results

##### Numerical results

##### Inferencial results

### Discussion and conclusion

